---
title: "Development actions history"
output: html_document
editor_options: 
  chunk_output_type: console
---

All commands that you use to use when developing packages...

# First time just after creating the project

- Fill the following chunk to create the DESCRIPTION of your package

```{r description}
# Describe your package
fusen::fill_description(
  pkg = here::here(),
  fields = list(
    Title = "Comparison of seasonal adjustment methods on ABS data",
    Version = "0.1",
    Description = "This package contains functions to download and clean time series data from the Australian Bureau of Statistics (ABS) and compare the outputs from the seasonal adjustment and trend methods of JDemetra+ and X-13ARIMA-SEATS against those published by the ABS (which uses SEASABS).",
    `Authors@R` = c(
      person("A", "Russo", email = "a.s.russo@outlook.com", role = c("aut", "cre"))
    )
  ),
  overwrite = TRUE
)
# Define License with use_*_license()
usethis::use_gpl_license()
```

# All-in-one function to deploy publicly on GitHub

Either use this GitHub all-in-one function or choose the steps one-by-one in the following sections.
See `vignette("share-on-a-github-website", package = "fusen")` for more information.

Not required since we Git is already being used

```{r, eval=FALSE}
# # _All-in-one share of your package and its documentation on GitHub
# fusen::init_share_on_github()
```


# Start using git

Not required since we Git is already being used

```{r, eval=FALSE}
# usethis::use_git()
# # Deal with classical files to ignore
# usethis::git_vaccinate()
# # Use main for primary branch
# usethis::git_default_branch_rename()
```

# Set extra sources of documentation

```{r, eval=FALSE}
# Install a first time
remotes::install_local()
# README
usethis::use_readme_rmd()
# Code of Conduct
usethis::use_code_of_conduct("contact@fake.com")
# NEWS
usethis::use_news_md()
```

**From now, you will need to "inflate" your package at least once to be able to use the following commands. Let's go to your flat template, and come back here later if/when needed.**


# Inflate flat_save_data.Rmd

Start by inflating `dev/flat_save_data.Rmd`:

```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_save_data.Rmd",
               vignette_name = "dev_save_data",
               overwrite = T)

```

# Include datasets

```{r}
local({
  knitr::knit("vignettes/dev_save_data.Rmd", output = tempfile())
  usethis::use_data(CPI, overwrite = TRUE)
  usethis::use_data(employment, overwrite = TRUE)
  usethis::use_data(GDP, overwrite = TRUE)
  usethis::use_data(inventories, overwrite = TRUE)
  usethis::use_data(retail, overwrite = TRUE)
  usethis::use_data(sales, overwrite = TRUE)
})

```

```{r}
fusen::inflate(flat_file = "dev/data_doc.Rmd",
               vignette_name = NA,
               overwrite = T)
```

# Package development tools
## Use once

```{r, eval=FALSE}
# Pipe
usethis::use_pipe()

# package-level documentation
usethis::use_package_doc()

# GitHub
# Add your credentials for GitHub
usethis::create_github_token()
gitcreds::gitcreds_set()

# _Or one step at a time
# Send your project to a new GitHub project (included in `init_share_on_github()`)
usethis::use_github()

# Set Continuous Integration
# _GitHub (included in `init_share_on_github()`)
usethis::use_github_action_check_standard()
usethis::use_github_action("pkgdown")
usethis::use_github_action("test-coverage")
# _GitLab
gitlabr::use_gitlab_ci(type = "check-coverage-pkgdown")
```

## Use everytime needed

```{r, eval=FALSE}
# Simulate package installation
pkgload::load_all()

# Generate documentation and deal with dependencies
attachment::att_amend_desc()

# Check the package
devtools::check()

# Add a new flat template
fusen::add_flat_template("add")
```

# Share the package

```{r, eval=FALSE}
# set and try pkgdown documentation website locally
usethis::use_pkgdown()
pkgdown::build_site()

# build the tar.gz with vignettes to share with others
devtools::build(vignettes = TRUE)

# Share your package and its documentation on GitHub
# usethis::create_github_token()
# gitcreds::gitcreds_set()
fusen::init_share_on_github()
```

